---
title: "You Draw It: Exponential Prediction"  
subtitle: "ISU Graphics Group"
author: "Emily Robinson"
date: "April 8, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	options(knitr.kable.NA = ''), 
	fig.height = 12, 
	fig.width = 12,
	fig.align = 'center'
)
library(RSQLite)
library(DBI)
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
```

## Discussion Points
+ Fix randomization? Not sure what is going on here but some people saw some plots twice and others never.
+ Simulate data across entire domain to fit line through. Seems like 0.5 and 0.75 differ in their estimated lines from `nls()`
    + Do we want the same data but truncated at different x-values? (e.g. x = 10 and x = 15).
+ How would we compare Sums of Squares over two different curvature levels (i.e. very different domains)?
+ What distribution should Sums of Squares follow? (*log transformation*, Gaussian with a log link, Gamma, Lognormal?)
+ Need a way to measure over / under estimation.

## Data

```{r obtain-data}
# Connect to database
filename <- "../you_draw_it_data.db"
sqlite.driver <- dbDriver("SQLite")
db_con <- dbConnect(sqlite.driver, dbname = filename)
# dbListTables(db_con)

# Import exponential prediction parameter details
exp_parameter_details <- dbReadTable(db_con,"exp_parameter_details") %>%
  select(parm_id, beta, sd, points_end_scale) %>%
  mutate(parm_id = as.factor(as.character(parm_id)),
         beta = as.factor(as.character(beta)),
         sd = as.factor(as.character(sd)),
         points_end_scale = as.factor(as.character(points_end_scale)))

# Import user data
users <- dbReadTable(db_con,"users") %>%
  select(-ip_address) %>%
  mutate(participantID = as.factor(as.character(as.numeric(as.factor(paste(nick_name, "_", study_starttime, sep = ""))))))

# Import exponential prediction feedback data
exp_feedback <- dbReadTable(db_con,"feedback") %>%
  filter(parm_id %in% c("exp_1", "exp_2", "exp_3", "exp_4")) %>%
  left_join(exp_parameter_details, by = "parm_id") %>%
  mutate(scale = ifelse(linear == "true", "Linear", "Log"),
         residual.drawn = ydrawn - y) %>%
  left_join(users, by = c("nick_name", "study_starttime")) %>%
  filter(recruitment == "Graphics Group")  %>%
  mutate(plotID = as.factor(as.character(as.numeric(as.factor(paste(nick_name, "_", start_time, "_", end_time, sep = ""))))),
         parm_id = as.factor(as.character(parm_id)),
         beta = as.factor(as.character(beta)),
         sd = as.factor(as.character(sd)),
         points_end_scale = as.factor(as.character(points_end_scale)))

# Import exponential prediction simulated data
exp_simulated_data <- dbReadTable(db_con,"simulated_data") %>%
  filter(parm_id %in% c("exp_1", "exp_2", "exp_3", "exp_4")) %>%
  left_join(exp_parameter_details, by = "parm_id") %>%
  left_join(users, by = c("nick_name", "study_starttime")) %>%
  filter(recruitment == "Graphics Group") %>%
  mutate(parm_id = as.factor(as.character(parm_id)),
         beta = as.factor(as.character(beta)),
         sd = as.factor(as.character(sd)),
         points_end_scale = as.factor(as.character(points_end_scale)))

# Disconnect from database
dbDisconnect(db_con)
```

### Feedback Data Summary

It appears some users saw some exponential parameter ID's twice and others never.

```{r feedback-summary}
exp_feedback %>%
  select(parm_id, beta, sd, points_end_scale, scale, participantID, plotID) %>%
  unique() %>%
  group_by(parm_id, beta, sd, points_end_scale, scale, participantID) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(participantID) %>%
  mutate(total = sum(count)) %>%
  ungroup() %>%
  mutate(plot = paste(parm_id, scale, sep = "; ")) %>%
  arrange(plot) %>%
  pivot_wider(id_cols = c("participantID", "total"),
              names_from = "plot",
              values_from = "count") %>%
  arrange(as.numeric(as.character(participantID))) %>%
  knitr::kable()
```

### Simulated Data Summary

But the data were simulated properly.

```{r simulated-summary}
exp_simulated_data %>%
  filter(dataset == "line_data", x == 20.00) %>%
  group_by(parm_id, beta, sd, points_end_scale, participantID) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(participantID) %>%
  mutate(total = sum(count)) %>%
  ungroup() %>%
  pivot_wider(id_cols = c("participantID", "total"),
              names_from = "parm_id",
              values_from = "count") %>%
  arrange(as.numeric(as.character(participantID))) %>%
  knitr::kable()
```

### Randomization Script

And I believe the randomization technique is correct.

```{r randomization-script, echo = T}
randomization_dataset <- tibble(parm_id = c(c("exp_1","exp_2","exp_3","exp_4"),c("exp_1","exp_2","exp_3","exp_4"), c("S","F","V","N")),
                                linear  = c(rep("true",4), rep("false",4), rep("true",4))
                                )
parm_ids <- sample(seq(1,12), 12, replace = F)

randomization_dataset %>% knitr::kable() 
parm_ids %>% t() %>% knitr::kable()
```

## Loess Smoother

```{r loess-smoother, echo = T}

# Fit Loess Smoother
loess.models <- exp_feedback %>%
        tidyr::nest(-plotID) %>%
        dplyr::mutate(
                # Perform loess calculation on each plotID
                loess.fit = purrr::map(data, loess,
                               formula = ydrawn ~ x),
                # Retrieve the fitted values from each model
                yloess = purrr::map(loess.fit, `[[`, "fitted")
        )

# Apply fitted y's as a new column
exp_feedback.smooth <- loess.models %>%
        dplyr::select(-loess.fit) %>%
        tidyr::unnest() %>%
  mutate(residual.loess = yloess - y) %>%
  select(participantID, age, gender, academic_study, recruitment, plotID, parm_id, scale, beta, sd, points_end_scale, x, y, ydrawn, yloess, residual.drawn, residual.loess, nick_name, study_starttime, start_time, end_time)

factorCols <- c("plotID", "parm_id", "nick_name", "study_starttime", "beta", "sd", "points_end_scale", "scale", "age", "gender", "academic_study", "recruitment", "participantID")
exp_feedback.smooth[,factorCols] <- lapply(exp_feedback.smooth[,factorCols], factor)

summary(exp_feedback.smooth)
```

## Spaghetti Plots

### Simulated Data

```{r spaghetti-simulated}
exp_simulated_data %>%
  filter(dataset == "line_data") %>%
  arrange(participantID, x) %>%
  ggplot(aes(x = x, y = y, group = participantID)) +
  geom_line(alpha = 0.3) +
  facet_grid(beta ~ points_end_scale, scales = "free", labeller = labeller(beta = label_both, points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 1)
```

### All Overlaid (No Smoother)

```{r spaghetti-all-overlaid-no-smoother}
exp_simulated_data %>%
  filter(dataset == "line_data") %>%
  arrange(participantID, x) %>%
  ggplot(aes(x = x)) +
  geom_line(data = exp_feedback.smooth, aes(y = ydrawn, group = plotID, color = scale), alpha = 0.7) +
  geom_line(alpha = 0.7, aes(y = y, group = participantID)) +
  facet_grid(beta ~ points_end_scale, scales = "free", labeller = labeller(beta = label_both, points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_color_manual("Scale", values = c("steelblue", "orange")) +
  scale_x_continuous(limits = c(10, 20))
```

### All Overlaid (Loess Smoother)

```{r spaghetti-all-overlaid-loess-smoother}
exp_simulated_data %>%
  filter(dataset == "line_data") %>%
  arrange(participantID, x) %>%
  ggplot(aes(x = x)) +
  geom_line(data = exp_feedback.smooth, aes(y = yloess, group = plotID, color = scale), alpha = 0.7) +
  geom_line(alpha = 0.2, aes(y = y, group = participantID)) +
  facet_grid(beta ~ points_end_scale, scales = "free", labeller = labeller(beta = label_both, points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_color_manual("Scale", values = c("steelblue", "orange")) +
  scale_x_continuous(limits = c(10, 20))
```

<hr>
### Residuals (No Smoother)

```{r residuals-no-smoother}
exp.spaghetti.plot.raw <- exp_feedback.smooth %>%
  arrange(participantID, parm_id, scale, x) %>%
  ggplot(aes(x = x, y = residual.drawn, group = plotID, color = scale)) +
  geom_line(alpha = 0.7) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", alpha = 0.5) +
  facet_grid(beta ~ points_end_scale, scales = "free", labeller = labeller(beta = label_both, points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_color_manual("Scale", values = c("steelblue", "orange"))
exp.spaghetti.plot.raw 
```

### Residual (Loess Smoother)

```{r residuals-loess-smoother}
exp.spaghetti.plot.smooth <- exp_feedback.smooth %>%
  arrange(participantID, parm_id, scale, x) %>%
  ggplot(aes(x = x, y = residual.loess, group = plotID, color = scale)) +
  geom_line(alpha = 0.7) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", alpha = 0.5) +
  facet_grid(beta ~ points_end_scale, scales = "free", labeller = labeller(beta = label_both, points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_color_manual("Scale", values = c("steelblue", "orange"))
exp.spaghetti.plot.smooth
```

## Sum of Squares

```{r SS-summary}
exp_SSdata <- exp_feedback.smooth %>%
  group_by(participantID, age, gender, academic_study, recruitment, plotID, parm_id, scale, beta, sd, points_end_scale) %>%
  summarise(SSdrawn = sum(residual.drawn^2),
            SSloess= sum((residual.loess)^2),
            count = n()) %>%
  ungroup()

exp_SSdata %>%
  ggplot(aes(x = scale, y = SSdrawn, color = participantID)) +
  geom_jitter(height = 0, width = 0.2, size = 2) +
  facet_grid(beta ~ points_end_scale, scales = "free", labeller = labeller(beta = label_both, points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_color_brewer(palette = "Paired") +
  scale_y_continuous("Sum of Squared \n Residuals (Drawn)") +
  scale_x_discrete("Scale")
```

### Beta = 0.1 (less curvature) (LMM with a log(SS) transformation)

<hr>
#### Drawn

```{r SS-drawn-01-lmm, fig.height = 6, fig.width = 12}
# SSdrawn.fit.01 <- glmer(log(SSdrawn) ~ scale*points_end_scale + (1|participantID),
                    # data = exp_SSdata %>% filter(beta == "0.1"),
                    # family = gaussian(link = "log"))
SSdrawn.fit.01 <- lmer(log(SSdrawn) ~ scale*points_end_scale + (1|participantID),
                    data = exp_SSdata %>% filter(beta == "0.1"))
anova(SSdrawn.fit.01) %>% knitr::kable(digits = 4)

SSdrawn.emmeans.01 <- emmeans(SSdrawn.fit.01, ~ scale:points_end_scale, type = "response")

SSdrawn.emmeans.01 %>% 
  as_tibble() %>%
  ggplot(aes(x = scale, y = response, fill = scale)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  # geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.2) +
  facet_grid(~points_end_scale, labeller = labeller(points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 0.75) +
  scale_fill_manual("Scale", values = c("steelblue", "orange")) 


SSdrawn.scale.01   <- emmeans(SSdrawn.fit.01, ~ scale | points_end_scale, type = "response")
pairs(SSdrawn.scale.01) %>% knitr::kable(digits = 4)


SSdrawn.points.01   <- emmeans(SSdrawn.fit.01, ~ points_end_scale | scale, type = "response")
pairs(SSdrawn.points.01) %>% knitr::kable(digits = 4)
```

<hr>
#### Smooth

```{r SS-smooth-01-lmm, fig.height = 6, fig.width = 12}
# SSloess.fit.01 <- glmer(log(SSloess) ~ scale*points_end_scale + (1|participantID),
                    # data = exp_SSdata %>% filter(beta == "0.1"),
                    # family = gaussian(link = "log"))
SSloess.fit.01 <- lmer(log(SSloess) ~ scale*points_end_scale + (1|participantID),
                    data = exp_SSdata %>% filter(beta == "0.1"))
anova(SSloess.fit.01) %>% knitr::kable(digits = 4)

SSloess.emmeans.01 <- emmeans(SSloess.fit.01, ~ scale:points_end_scale, type = "response")

SSloess.emmeans.01 %>% 
  as_tibble() %>%
  ggplot(aes(x = scale, y = response, fill = scale)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  # geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.2) +
  facet_grid(~points_end_scale, labeller = labeller(points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 0.75) +
  scale_fill_manual("Scale", values = c("steelblue", "orange")) 


SSloess.scale.01   <- emmeans(SSloess.fit.01, ~ scale | points_end_scale, type = "response")
pairs(SSloess.scale.01) %>% knitr::kable(digits = 4)


SSloess.points.01   <- emmeans(SSloess.fit.01, ~ points_end_scale | scale, type = "response")
pairs(SSloess.points.01) %>% knitr::kable(digits = 4)
```

### Beta = 0.23 (more curvature) (LMM with a log(SS) transformation)

<hr>
#### Drawn

```{r SS-drawn-23-glmm, fig.height = 6, fig.width = 12}
# SSdrawn.fit.23 <- glmer(log(SSdrawn) ~ scale*points_end_scale + (1|participantID),
#                     data = exp_SSdata %>% filter(beta == "0.23"),
#                     family = gaussian(link = "log"))
SSdrawn.fit.23 <- lmer(log(SSdrawn) ~ scale*points_end_scale + (1|participantID),
                    data = exp_SSdata %>% filter(beta == "0.23"))
anova(SSdrawn.fit.23) %>% knitr::kable(digits = 4)

SSdrawn.emmeans.23 <- emmeans(SSdrawn.fit.23, ~ scale:points_end_scale, type = "response")

SSdrawn.emmeans.23 %>% 
  as_tibble() %>%
  ggplot(aes(x = scale, y = response, fill = scale)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  # geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.2) +
  facet_grid(~points_end_scale, labeller = labeller(points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 0.75) +
  scale_fill_manual("Scale", values = c("steelblue", "orange")) 


SSdrawn.scale.23   <- emmeans(SSdrawn.fit.23, ~ scale | points_end_scale, type = "response")
pairs(SSdrawn.scale.23) %>% knitr::kable(digits = 4)


SSdrawn.points.23   <- emmeans(SSdrawn.fit.23, ~ points_end_scale | scale, type = "response")
pairs(SSdrawn.points.23) %>% knitr::kable(digits = 4)
```

<hr>
#### Smooth

```{r SS-loess-23-glmm, fig.height = 6, fig.width = 12}
# SSloess.fit.23 <- glmer(log(SSloess) ~ scale*points_end_scale + (1|participantID),
#                     data = exp_SSdata %>% filter(beta == "0.23"),
#                     family = gaussian(link = "log"))
SSloess.fit.23 <- lmer(log(SSloess) ~ scale*points_end_scale + (1|participantID),
                    data = exp_SSdata %>% filter(beta == "0.23"))
anova(SSloess.fit.23) %>% knitr::kable(digits = 4)

SSloess.emmeans.23 <- emmeans(SSloess.fit.23, ~ scale:points_end_scale, type = "response")

SSloess.emmeans.23 %>% 
  as_tibble() %>%
  ggplot(aes(x = scale, y = response, fill = scale)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  # geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.2) +
  facet_grid(~points_end_scale, labeller = labeller(points_end_scale = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 0.75) +
  scale_fill_manual("Scale", values = c("steelblue", "orange")) 


SSloess.scale.23   <- emmeans(SSloess.fit.23, ~ scale | points_end_scale, type = "response")
pairs(SSloess.scale.23) %>% knitr::kable(digits = 4)


SSloess.points.23   <- emmeans(SSloess.fit.23, ~ points_end_scale | scale, type = "response")
pairs(SSloess.points.23) %>% knitr::kable(digits = 4)
```

