---
title: "Perception (Lineups)"
author: "Emily Robinson"
date: "Spring 2022"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
library(tidyverse)
library(cowplot)
library(digest)

library(RSQLite)
library(DBI)
library(here)
library(readr)

library(lme4)
library(emmeans)
```

# Data

```{r, eval = F}

# user data
db_con <- dbConnect(dbDriver("SQLite"), dbname = here("perception-of-statistical-graphics/databases/00_demographics_db.db"))

users <- dbReadTable(db_con,"users") %>%
  rowwise() %>%
  mutate(participant_id = digest(prolific_id, "md5"))

completed_sections_orig <- dbReadTable(db_con,"completed_sections") %>%
  rowwise() %>%
  mutate(participant_id = digest(prolific_id, "md5"))

dbDisconnect(db_con)

completed_lineups <- completed_sections_orig %>%
  filter(section_complete == "lineups")

completed_sections_all <- completed_sections_orig %>%
  pivot_wider(id_cols = c("nick_name", "ip_address", "study_starttime"),
              names_from = "section_complete",
              values_from = time)

demographic_data <- full_join(users, completed_sections_all, by = c("nick_name", "ip_address", "study_starttime"))  %>%
  rename(you_draw_it = `you-draw-it`) %>%
  mutate(lineups = ifelse(is.null(lineups), NA, lineups),
         you_draw_it = ifelse(is.null(you_draw_it), NA, you_draw_it),
         estimation = ifelse(is.null(estimation), NA, estimation)) %>%
  mutate(studies_complete = as.numeric(!is.na(lineups)) + as.numeric(!is.na(you_draw_it)) + as.numeric(!is.na(estimation))) %>%
  arrange(-studies_complete)

all_participants <- demographic_data %>% filter(studies_complete == 3, recruitment != "I am the researcher") %>%
  select(participant_id) %>%
  as.vector()
lineup_participants <- demographic_data %>% filter(studies_complete == 1, recruitment != "I am the researcher")

# lineup data
db_con <- dbConnect(dbDriver("SQLite"), dbname = here("perception-of-statistical-graphics/databases/01_lineups_db.db"))

lineups_feedback <- dbReadTable(db_con,"feedback")  %>%
  rowwise() %>%
  mutate(participant_id = digest(prolific_id, "md5"))

picture_details <- dbReadTable(db_con,"picture_details")

dbDisconnect(db_con)

# combine lineup feedback and user data
joinCols <- c("participant_id", "nick_name", "ip_address", "study_starttime", "prolific_id")
lineup_data_orig <- users %>%
  full_join(completed_lineups, by = c("participant_id", "study_starttime", "nick_name", "ip_address", "prolific_id")) %>%
  filter(recruitment != "I am the researcher") %>%
  right_join(lineups_feedback, by = joinCols)

factorCols <- c("participant_id", "nick_name", "ip_address", "prolific_id", "age", "gender", "academic_study", "computer_mouse", "recruitment", "pic_id", "param_value", "set", "test_param", "conf_level")
lineup_data_orig[,factorCols] <- lapply(lineup_data_orig[,factorCols], factor)

lineup_model_data <- lineup_data_orig %>%
  separate(param_value, into = c(NA, "target", NA, "null", NA), sep = "-") %>%
  filter(target != null) %>%
  mutate(curvature = paste0("t-", target, "_n-", null), .after = null) %>%
  rowwise() %>%
  mutate(dataset_id = as.factor(digest(paste(curvature, set, sep = "-"), "md5")), .after = set) %>%
  rename(scale = test_param) %>%
  mutate(binary_response = ifelse(correct %in% as.numeric(unlist(strsplit(response_no, split = ","))), 1, 0), .after = response_no) %>%
  group_by(participant_id, nick_name, ip_address, study_starttime, prolific_id) %>%
  mutate(n_lineups = n()) %>%
  ungroup() %>%
  filter(!is.na(section_complete)) %>%
  # filter(participant_id %in% all_participants$participant_id) %>%
  select(participant_id, n_lineups, age, gender, academic_study, computer_mouse, recruitment, dataset_id, set, curvature, target, null, scale, correct, response_no, binary_response, conf_level, choice_reason)

write.csv(lineup_model_data, here("analyses/01-lineups/lineup-model-data.csv"), row.names = F, na = "")
```

```{r, eval = T}
db_con <- dbConnect(dbDriver("SQLite"), dbname = here("perception-of-statistical-graphics/databases/01_lineups_db.db"))
lineups_feedback <- dbReadTable(db_con,"feedback")  %>%
  rowwise() %>%
  mutate(participant_id = digest(prolific_id, "md5"))
dbDisconnect(db_con)

lineup_model_data <- read.csv(here("analyses/01-lineups/lineup-model-data.csv"))
```

```{r}
lineup_model_data %>%
  select(participant_id, age, gender, academic_study, dataset_id, curvature, scale, correct, response_no, binary_response, conf_level) %>%
  head() %>%
  knitr::kable()
```

<!-- # Check Randomization -->

```{r, eval = F}
lineup_model_data %>%
  group_by(participant_id, n_lineups, curvature, scale) %>%
  summarize(n = n()) %>%
  filter(scale == "linear") %>%
  pivot_wider(id_cols = c("participant_id", "n_lineups"),
              names_from = "curvature",
              values_from = "n") %>%
  arrange(-n_lineups) %>%
  knitr::kable(caption = "Linear Scale")

lineup_model_data %>%
  group_by(participant_id, n_lineups, curvature, scale) %>%
  summarize(n = n()) %>%
  filter(scale == "log") %>%
  pivot_wider(id_cols = c("participant_id", "n_lineups"),
              names_from = "curvature",
              values_from = "n") %>%
  arrange(-n_lineups) %>%
  knitr::kable(caption = "Log Scale")
```

# Analysis

```{r}
ploteval_summary <- lineup_model_data %>% group_by(curvature, set, scale) %>% summarize(count = n()) %>% arrange(count)
```

+ There are `r length(unique(lineups_feedback$participant_id)) %>% as.numeric` participants and `r nrow(lineups_feedback) %>% as.numeric()` lineups complete. 
+ There are `r length(unique(lineup_model_data$participant_id)) %>% as.numeric` participants and `r nrow(lineup_model_data) %>% as.numeric()` lineups complete. 
+ Each plot was evaluated between `r min(ploteval_summary$count) %>% as.numeric() %>% round(2)` and `r max(ploteval_summary$count)  %>% as.numeric()  %>% round(2)` times (Mean: `r mean(ploteval_summary$count)  %>% as.numeric()  %>% round(2)`, SD: `r sd(ploteval_summary$count)  %>% as.numeric()  %>% round(2)`).
+ Participants correctly identified the target panel in `r round(as.numeric(mean(lineup_model_data$binary_response[lineup_model_data$scale == 'linear']))*100, 2)`% of the `r nrow(lineup_model_data[lineup_model_data$scale == 'linear',]) %>% as.numeric()` lineup evaulations made on the linear scale and `r round(as.numeric(mean(lineup_model_data$binary_response[lineup_model_data$scale == 'log']))*100, 2)`% of the `r nrow(lineup_model_data[lineup_model_data$scale == 'log',]) %>% as.numeric()` lineup evaluations made on the log scale.

Target plot identification was analyzed using the `glmer` function in the `lme4` R package. 
Estimates and odds ratio comparisons between the log and linear scales were calculated using the `emmeans` R package.
Each lineup plot evaluated was assigned a value based on the participant response (correct = 1, not correct = 0).
Define $Y_{ijkl}$ to be the event that participant $l = 1,...,N_{participant}$ correctly identifies the target plot for data set $k = 1,2$ with curvature combination $j = 1,2,3,4,5,6$ plotted on scale $i = 1,2$.
The binary response was analyzed using a generalized linear mixed model following a binomial distribution with a logit link function with a row-column blocking design accounting for the variation due to participant and data set respectively as 
\begin{equation}
\text{logit }P(Y_{ijk}) = \eta + \delta_i + \gamma_j + \delta \gamma_{ij} + s_l + d_k
\end{equation}
\noindent where

+ $\eta$ is the baseline average probability of selecting the target plot
+ $\delta_i$ is the effect of scale $i = 1,2$
+ $\gamma_j$ is the effect of curvature combination $j = 1,2,3,4,5,6$
+ $\delta\gamma_{ij}$ is the two-way interaction between the $i^{th}$ scale and $j^{th}$ curvature combination
+ $s_l \sim N(0,\sigma^2_\text{participant})$is the random effect for participant characteristics
+ $d_k \sim N(0,\sigma^2_{\text{data}})$ is the random effect for data specific characteristics. 

\noindent We assume that random effects for data set and participant are independent.

```{r}
glmm_mod <- glmer(binary_response ~ curvature*scale + 
                            (1 | participant_id) +
                            (1 | dataset_id),
                  data = lineup_model_data,
                  family = binomial(link = "logit"))

lineup_anova <- car::Anova(glmm_mod) %>% broom::tidy()
summary(glmm_mod)

# write.csv(lineup_anova, here("analyses/01-lineups/results/lineups-anova.csv"), row.names = F, na =)

lineup_anova %>% knitr::kable(digits = 3, caption = "Lineups GLMM ANOVA")
```

```{r, eval = F, include = F, fig.align = 'center', fig.width = 12, fig.height = 6}
lineup_lsmeans <- emmeans(glmm_mod, specs = ~ "curvature:scale", type = "response") %>%
  as_tibble()

lineup_lsmeans %>%
  mutate(curvature = factor(curvature, levels = c("t-E_n-H", "t-H_n-E", "t-E_n-M", "t-M_n-E", "t-M_n-H", "t-H_n-M"))) %>%
  ggplot(aes(x = curvature, y = prob, group = scale)) +
  geom_bar(stat = "identity", fill = "lightgray") +
  # geom_text(aes(y = asymp.UCL + 0.05)) +
  facet_wrap(~scale, ncol = 1) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  theme_bw() +
  scale_y_continuous("Probability of target panel detected", limit = c(-0.05,1.1), breaks = seq(0,1,0.2)) +
  scale_x_discrete("Curvature")
```


```{r, fig.align = 'center', fig.width = 12, fig.height = 6}
lineups_or <- emmeans(glmm_mod, ~ scale | curvature, type = "response") %>% pairs(reverse = T, infer = c(T,T)) %>% as_tibble()
write.csv(lineups_or, here("analyses/01-lineups/results/lineups-odds-ratios.csv"), row.names = F, na =)

dodge <- position_dodge(width=0.9)
lineups_or_plot <- lineups_or %>%
  separate(curvature, into = c(NA, "target", NA, "null")) %>%
  mutate(null = factor(null, levels = c("E", "M", "H"), labels = c("High Curvature", "Medium Curvature", "Low Curvature")),
         target = factor(target, levels = c("E", "M", "H"), labels = c("High Curvature", "Medium Curvature", "Low Curvature"))) %>%
  ggplot(aes(x = odds.ratio, y = null, color = target, shape = target)) + 
  geom_point(position = dodge, size = 3) + 
  geom_errorbar(aes(xmin = asymp.LCL, xmax = asymp.UCL), position = dodge, width = .1) +
  geom_vline(xintercept = 1) +
  theme_bw()  +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        legend.key.size = unit(1, "line"),
        legend.position = "bottom"
  ) +
  scale_y_discrete("Null Panel Difficulty", position = "left") +
  scale_x_continuous("Odds ratio (on log scale) \n (Log vs Linear)", trans = "log10", labels = scales::comma) + 
  scale_color_manual("Target Panel Difficulty", values = c("#004400", "#116611", "#55aa55")) + 
  scale_shape_discrete("Target Panel Difficulty")

picsList <- c("images/tM_nH_thumbnail.png", 
              "images/tE_nH_thumbnail.png", 
              "images/tH_nM_thumbnail.png", 
              "images/tE_nM_thumbnail.png", 
              "images/tH_nE_thumbnail.png", 
              "images/tM_nE_thumbnail.png"
)

pimage <- axis_canvas(lineups_or_plot, axis = 'y') + 
  draw_image(picsList[1], y = 2.75, scale = 0.7) +
  draw_image(picsList[2], y = 2.25, scale = 0.7) +
  draw_image(picsList[3], y = 1.75, scale = 0.7) +
  draw_image(picsList[4], y = 1.25, scale = 0.7) +
  draw_image(picsList[5], y = 0.75, scale = 0.7) +
  draw_image(picsList[6], y = 0.25, scale = 0.7)

# insert the image strip into the plot
ggdraw(insert_yaxis_grob(lineups_or_plot, pimage, position = "right", clip = "on"))
```



